---
title: "Candidate Names"
---

### Creating a new csv file with a tally for each time the name occurs in the data

```{r}

library(tidyverse)

#Data source: democracy club, candidate list, released 07/06/2024

name <- read.csv("candidate_name/candidate_GE24.csv")

```

```{r}

#replace captain beany with name

name$person_name <- sub("Captain Beany", "Barry Kirk", name$person_name, fixed = TRUE)

name$person_name <- sub("The Zok", "Zoran Zoc", name$person_name, fixed = TRUE)

#tally first names

first_name_tally <- name %>%
  mutate(cleaned_names = gsub("Sir |Lord |Doctor |Dr |M. ", "", person_name)) %>% 
  mutate(first_name = sub(" .*", "", cleaned_names)) %>%  
  count(first_name, sort = TRUE)  

#create a new column with all lower case

first_name_tally <- first_name_tally %>%
  mutate(name_lower = tolower(first_name))

# anomalies in dataset: monster raving loony party candidate names, AI steve, emperor of india prince, Wing Commander (RTD) Fred Ramsey

```

### Assigning names: method

The following method was used to assign gender to names. After the full candidate list had been downloaded from the Democracy Club, a tally of names were summarised to create a list of 1,374 unique names. Lists were created to match gender to names. Clearly gendered names such as 'David' and 'Sarah' were automatically put into their corresponding list. Gender-neutral names were left out to be checked on a case by case basis.

However, once the list of 1,374 names had reached the point where only one instance of each name occurred, the names were checked on a case by case basis and assigned to the corresponding list along with the gendered names. So, for example, although 'Rory' and 'Robyn' are gender neutral names, they appear in the male or female name lists as only one candidate had this name and their gender was verified specifically. To check, names were google searched to find a picture and ideally a pronoun reference to the candidate (in, say, a journal article about them) to verify their gender. If their gender was unverifiable, or it appeared they did not identify with the gender binary, they were placed into a separate list.

This is not a perfect method, and aimed only at achieving a general sense of proportions between men and women candidates within each party. Mistakes may have been made in individual cases.

```{r}

name_tally <- name %>%
  filter(grepl("^Hilda ", person_name)) %>%  
  group_by(party_name) %>%  # Group by party_name
  summarise(count = n())  # Count occurrences


# split up first name and last name into new column

name <- name %>%
  select(person_name, party_name, post_label) %>%
  mutate(cleaned_names = gsub("Sir |Lord |Doctor |Dr |M. ", "", person_name)) %>% 
  mutate(first_name = sub(" .*", "", cleaned_names),
         last_name = str_extract(cleaned_names, "\\w+$")) %>%
  select(person_name, first_name, last_name, party_name, post_label)

# Combine Labour and the cooperative party into 1


# male, female and unknown names matched

male_names1 <- c("David", "John", "Paul", "James", "Mark", "Andrew", "Richard", "Michael", "Ian", "Stephen", "Steve", "Peter", "Simon", "Robert", "Tom", "Ben", "Jonathan", "Andy", "Matthew", "Mike", "Martin", "Neil", "Tony", "Dan", "Daniel", "Christopher", "Jack", "Matt", "Tim", "Adam", "Graham", "Alan", "Craig", "Luke", "Edward", "Gary", "Nigel")

male_names2 <- c("Gareth", "Darren", "Scott", "Stuart", "Thomas", "Oliver", "Rob", "Bill", "Jason", "Kevin", "Nicholas", "Dave", "George", "Jim", "Josh", "Keith", "Mohammed", "Barry", "Jeremy", "Brian", "Charles", "Colin", "Philip", "Alexander", "Anthony", "Bob", "Ed", "Guy", "Henry", "Pete", "Phil", "Roger", "Will", "William", "Jon", "Liam", "Patrick", "Stewart", "Shaun", "Trevor", "Alastair", "Dominic", "Frank", "Gavin", "Iain", "Jacob", "Jake", "Jeff", "Karl", "Malcolm", "Steven", "Adrian", "Damian", "Joshua", "Khalid", "Kieran", "Marcus", "Alistair", "Callum", "Carl", "Clive", "Declan", "Donald", "Duncan", "Gordon", "Grant", "Greg", "Hamish", "Harry", "Lewis", "Martyn", "Owen", "Samuel", "Tommy", "Benjamin", "Brandon", "Brendan", "Connor", "Derek", "Geoff", "Gregory", "Joseph")

male_names3 <- c("Marc", "Raj", "Toby", "Abdul", "Arnold", "Bernard", "Bradley", "Douglas", "Dylan", "Eddie", "Freddie", "Garry", "Glen", "Imran", "Johnny", "Justin", "Keir", "Kenneth", "Laurence", "Leo", "Mick", "Muhammad", "Nathan", "Niall", "Robbie", "Rupert", "Timothy", "Amir", "Angus", "Antonio", "Arthur", "Bobby", "Brent", "Brett", "Conor", "Conrad", "Damien", "Darryl", "Dean", "Desmond", "Doug", "Earl", "Edmund", "Edwin", "Eric", "Ethan", "Geoffrey", "Gerald", "Glenn", "Grahame", "Howard", "Hugo", "Joel", "Jordan", "Ken", "Khalil", "Kieron", "Kyle", "Lawrence", "Louie", "Mathew", "Mohamed", "Mohammad", "Muhammed", "Naveed", "Norman", "Oscar", "Phillip", "Raymond", "Rich", "Rod", "Ron", "Roy", "Russ", "Seamus", "Sebastian", "Shane", "Theo", "Victor", "Wayne", "Wesley", "Ahmed", "Aidan", "Alasdair", "Alberto", "Alfred", "A.", "AI", "Aadil", "Abbas", "Abdoul", "Abdulla", "Abul", "Ade", "Adedotun", "Afzal", "Ahmet", "Ahsan", "Ajaz", "Ajmal", "Akeel", "Akhlaque", "Akhmed", "Al")

male_names4 <- c("Aled", "Alexi", "AliRom", "Allister", "Altaf", "Alun", "Alyn", "Ameet", "Amjad", "Ammar", "Amrit", "Andre", "Andrei", "Anil", "Anwarul", "Archibald", "Arnie", "Arran", "Arron", "Asa", "Ashvir", "Athar", "Atif", "Attic", "Attiq", "Aubrey", "Audel", "Augustine", "Austin", "Ayoub", "Azhar", "Azmat", "Babar", "Baggy", "Bambos", "Barmy", "Barney", "Baron", "Bart", "Bastôn", "Bayo", "Benedict", "Benet", "Bim", "Blair", "Blaise", "Blake", "Brad", "Brenden", "Bret", "Brook", "Bruce", "Byron", "Cade", "Cahal", "Calum", "Calvin", "Cam", "Carlo", "Carne", "Cas", "Cathal", "Caz", "Chan", "Charith", "Chas", "Chaz", "Ciaran", "Citizen", "Clayton", "Cliff", "Cole", "Collan", "Colum", "Count", "Cris", "Crispian", "Dafydd", "Dai", "Dale", "Damon", "Darshan", "Daud", "Davinder", "Dawud", "Debojyoti", "Denis", "Dennis", "Des", "Dharmanand", "Dhruv", "Dhruva", "Dick", "Dickie", "Dillan", "Dino", "Dipesh", "Dom", "Dominick", "Donal", "Donnie", "Draeyk", "Drummond", "Dáire", "Ehtashamul", "Ejel", "Elliott", "Ellis", "Emmett", "Emperor", "Eraj", "Erik", "Ertan", "Eugene", "Ewan", "Ezechiel", "Fabian", "Fabio", "Faisal", "Faraz", "Farooq", "Fergal", "Fesl", "Filson", "Franck", "Fred")

male_names5 <- c("Frederick", "Gagan", "Garreth", "Ged", "Gerard", "Gez", "Ghazi", "Giane", "Gideon", "Giles", "Giuseppe", "Glyn", "Gregor", "Grumpus", "Gulvinder", "Gurinder", "Gus", "Gwenton", "Gwyn", "Habib", "Hamilton", "Hans", "Harbant", "Hari", "Haroun", "Harrison", "Hassan", "Heiko", "Helmi", "Hersh", "Hoo-Ray", "Horatio", "Howling", "Huge", "Iconic", "Iddrisu", "Ieuan", "Iftikhar", "Imad", "Imtiaz", "Ioan", "Iolo", "Iqbal", "Irakli", "Irfan", "Irwen", "Isaac", "Ishtiaq", "Issan", "Ivan", "Jackson", "Jag", "Jaginder", "Jahir", "Jamal", "Jamie-Lee", "Janus", "Javed", "Jed", "Jeevun", "Jeffrey", "Jehad", "Jenner", "Jerome", "Jerry", "Jesse", "Job", "Jody", "Johnathan", "Johno", "Jonathon", "Jonny", "Josiah", "Jovan", "Kabeer", "Kaine", "Kamel", "Kamran", "Kane", "Kanishka", "Kashif", "Kayode", "Kaz", "Kefentse", "Keiron", "Kelvin", "Ketankumar", "Kevan", "Khandu", "Kieren", "King", "Kiran", "Kit", "Knigel", "Korin", "Kuldev", "Kwaku", "Kyn", "Lalit", "Lance", "Larry", "Lars", "Laurent", "Laurie", "Lenny", "Leslie", "Lester", "Lincoln", "Lionel", "Lloyd", "Lorenzo", "Louis", "Lucas", "Lucian", "Luigi", "Lynden", "Maciej", "Mal", "Malachy", "Mamun", "Manny", "Manu", "Maqsood", "Marco", "Markus", "Mattie", "Matty", "Maurice", "Maxwell", "Mayuran", "Mehmood", "Merv", "Mian", "Michal", "Miles", "Minhajul", "Mitchell", "Mocky", "Mohhamed", "Monty", "Moui-Tabrizy", "Mustafa", "Mustaque", "Muzafar", "Myles", "Nagarjun", "Nagender", "Nahim", "Najmul", "Narinder", "Nasri", "Navdeep", "Navendu", "Navid", "Neale", "Neill", "Neville", "Niel", "Nimal", "Nisar", "Nizam", "Noah", "Nurul", "Obi", "Olivio", "Olli", "Ollie", "Olusola", "Omar")

male_names6 <- c("Osman", "Otto", "Owain", "Owais", "Paco", "Padraic", "Parham", "Parmjit", "Pedro", "Pelé", "Perran", "Pierce", "Piers", "Prabhdeep", "Prad", "Praful", "Primesh", "Priyank", "Psychobilly", "Qazi", "Rahoul", "Raja", "Rajesh", "Rajiv", "Ralph", "Ranil", "Ravaldeep", "Ravi", "Reginald", "Rehman", "Reza", "Rhodri", "Riad", "Rio", "Rishi", "Rizvi", "Roberto", "Roderick", "Rodney", "Roh", "Ronnie", "Rory", "Rowland", "Ruairi", "Ruraidh", "Ryk", "Saba", "Sadat", "Saiham", "Samsuzzaman", "Sandesh", "Saqib", "Sarajulhaq", "Satbir", "Sathish", "Sayeed", "Senthil", "Seyi", "Shafaei", "Shafat", "Shaffaq", "Shailesh", "Shakeel", "Shakir", "Sham", "Shangara", "Shasha", "Shaz", "Shazu", "Sheikh", "Sherief", "Shesh", "Shockat", "Sid", "Sidney", "Sojan", "Somon", "Spencer", "Stan", "Stanley", "Steffan", "Stevan", "Steward", "Sukhinder", "Suneil", "Surjit", "Sushil", "Séamas", "Tad", "Talat", "Tam", "Tan", "Tarek", "Tarik", "Tariq", "Tarun", "Tasos", "Tauqeer", "Tayab", "Teck", "Ted", "Theodore", "Thom", "Tiger", "Titus", "Tobias", "Tobie", "Todd", "Toqueer", "Torcuil", "Torsten", "Tosh", "Tristram", "Uday", "Udo", "Ukonu", "Ulysse", "Umair", "Umar", "Vinay", "Vinayak", "Vipul", "Vivek", "Vusi", "Wael", "Waheed", "Wais", "Wajad", "Wajahat", "Warinder", "Waseem", "Wes", "Wil", "Wing", "Wisdom", "Yas", "Yasin", "Yemi", "Yoel", "Yousaff", "Yousef", "Yousuf", "Zack", "Zaf", "Zaffar", "Zafran", "Zaid", "Zain", "Zak", "Zbigniew", "Zen", "Zhafaran", "Zohaib", "Zoran", "Zubir", "Zuffar", "Zulfikar", "Sadik", "Niko", "Adnan", "Hanif", "Amin", "Asif", "Antony", "Sean", "Allan", "Tomos", "Julian", "Alec", "Terence", "Fraser", "Olly", "Leon", "Aftab", "Felix", "Anand", "Ross", "Euan", "Don", "Syed", "Golam", "Vikas", "Elwyn", "Billy", "Nick", "Stefan", "Kenny", "Zahid", "Jay")

#FEMALE

female_names1 <- c("Sarah", "Helen", "Catherine", "Rachel", "Emma", "Claire", "Caroline", "Kate", "Laura", "Rebecca", "Lucy", "Amanda", "Jane", "Anna", "Hannah", "Julie", "Ruth", "Elizabeth", "Louise", "Karen", "Alison", "Angela", "Anne", "Jennifer", "Liz", "Maria", "Emily", "Joanna", "Lisa", "Sally", "Charlotte", "Christine", "Clare", "Debbie", "Julia", "Mary", "Susan", "Ann", "Jan", "Kim", "Lauren", "Pippa", "Rosie", "Vicky", "Victoria", "Wendy", "Cat", "Diana", "Elaine", "Fiona", "Jessica", "Katherine", "Katie", "Linda", "Lorraine", "Lynn", "Maureen", "Melanie", "Natalie", "Alexandra", "Amy", "Andrea", "Becca", "Carol", "Dawn", "Denise", "Ellie", "Gillian", "Heather", "Hilary", "Jayne", "Jill", "Kelly", "Lorna", "Michelle", "Pamela", "Patricia", "Samantha", "Sandra", "Sara","Sophie", "Sue", "Teresa", "Vanessa", "Yvonne", "Alice", "Amelia", "Angie", "Bridget", "Carla", "Christina", "Donna", "Georgie", "Helena", "Holly", "Jackie", "Jade", "Janet", "Janice", "Jenny", "Joy", "Katharine", "Kirsty", "Lara", "Margaret", "Marie", "Maxine", "Nadia", "Nancy", "Natasha", "Nicola", "Pauline", "Rachael", "Stephanie", "Vicki", "Yasmin", "Aisha", "Anita", "Anne-Marie", "Anneliese", "Annie", "Becky", "Caitlin")

female_names2 <- c("Cara", "Carolyn", "Celia", "Christie", "Danielle", "Deborah", "Deirdre", "Diane", "Eleanor", "Erin", "Esther", "Eva", "Fay", "Fionna", "Gemma", "Georgia", "Georgina", "Gina", "Gloria", "Harriet", "Hazel", "Imogen", "Irene", "Isobel", "Izzy", "Jacqueline", "Jemma", "Jen", "Judith", "Judy", "June", "Kathleen", "Kathryn", "Kayleigh", "Kiera", "Leanne", "Lizzie", "Lois", "Lyn", "Lynda", "Maggie", "Marianne", "Maya", "Mona", "Monica", "Olivia", "Paula", "Penelope", "Penny", "Phoebe", "Rowena", "Sharon", "Sheila", "Shona", "Siobhan", "Sonia", "Sophia", "Stella", "Suzanne", "Tessa", "Thérèse", "Zoe", "Aarti", "Abigail", "Adele", "Alicia", "Alixandra", "Allison", "Aasha", "Aasiya", "Abena", "Abi", "Abtisam", "Adria", "Aheesha", "Aishat", "Albati", "Amie", "Ammi", "Ange", "Angeliki", "Aysha", "Ani", "Anja", "Anna-Maria", "Annaliese", "Annette", "Antoinette", "Antonia", "Anum", "Aphra", "April", "Apsana", "Aroma", "Aruhan", "Asha", "Ayesha", "Barbara", "Beatrice", "Beccy", "Bell", "Bella", "Bernadette", "Beth", "Bianca", "Brenda", "Cadewyn", "Caitriona", "Candice", "Carissma", "Carline", "Carolyne", "Carolynn", "Caron", "Cassi", "Cassie", "Cate", "Cath", "Cathy", "Chantal", "Charley", "Charmaine", "Chelsea", "Chelsey", "Cheney", "Cherilyn", "Cheryl", "Cheryl-Ann", "Chi", "Chika", "Chipiliro", "Chloe", "Chloe-Jane", "Chloe-Louise", "Chrishni", "Clareine", "Claudia", "Colette", "Cordelia", "Corinthia", "Corri", "Daisy", "Damola", "Danni", "Danuta", "Deanna", "Debra", "Deidre", "Desi", "Devenne", "Devina", "Dine", "Dionne", "Dominie", "Déirdre", "Eishar", "Ekua", "Ellen", "Elsie", "Emma-Jane", "Emmie", "Ese", "Esme")

female_names3 <- c("Eunice", "Evelyn", "Faiza", "Farzana", "Faten", "Faye", "Felicity", "Fennie", "Feryal", "Fleur", "Flick", "Flo", "Florence", "Francesca", "Francoise", "Freya", "Gabrielle", "Gail", "Gayathri", "Gemma-Jane", "Gen", "Geraldine", "Gerri", "Giovanna", "Gurpreet", "Hajira", "Halima", "Hanna", "Harpreet", "Harriett", "Hayley", "Heidi", "Hema", "Hillary", "Hina", "Holly-Mae", "Irenea", "Isabel", "Isabella", "Isabelle", "Issy", "Ivana", "Jabu", "Jacqui", "Jamila", "Janey", "Jaswinder", "Javeria", "Jeannine", "Jenifer", "Jessie", "Jilleane", "Joan", "Joani", "Joanne", "Jodie", "Johanna", "Josie", "Ju", "Juliet", "Justine", "Kamala", "Kamla", "Karin", "Kathrine", "Kathy", "Katrina", "Katy", "Kelda", "Kellie-Jay", "Kemi", "Khalila", "Khobi", "Kirith", "Kirsteen", "Kirsten", "Kristy", "Kulsum", "Lady", "Laura-Beth", "Laura-Louise", "Layla", "Leena", "Leila", "Lex", "Leyla", "Leza", "Lia", "Lilian", "Lillian", "Linsey", "Lizzi", "Llinos", "Lola", "Lynne", "Maahwish", "Maddison", "Magdaline", "Mahtab", "Malihah", "Manda", "Mandy", "Manuela", "Marian", "Marianna", "Marie-Clair", "Marilyn", "Marion", "Marisa", "Marsha", "Martha", "Maryse", "Mehmoona", "Mehreen", "Mhairi", "Miatta", "Mims", "Miranda", "Miriam", "Moira", "Momtaz", "Moni", "Morgan", "Munira", "Muryam", "Nada", "Nanda", "Nandita", "Naomi", "Naranee", "Nariese", "Natalia", "Nataly", "Nathalie", "Naushabah", "Neena", "Nesil", "Netty", "Nia", "Niamh", "Nicky", "Nida", "Noor", "Nora", "Nuala", "Nupur", "Nus")

female_names4 <- c("Oana", "Onike", "Oyebola", "Pallavi", "Pam", "Paras", "Paulette", "Pavitar", "Peggy", "Perveen", "Petra", "Philippa", "Pinder", "Polly", "Poppy", "Preet", "Priti", "Rabina", "Reetendra", "Rehiana", "Reva", "Rhea", "Rhiannon", "Ria", "Rizwana", "Robyn", "Rochelle", "Roisin", "Rosalind", "Rosalinda", "Rose", "Rosemary", "Rosena", "Rosey", "Roxanne", "Roz", "Ruby", "Rufia", "Rumi", "Rupa", "Rushanara", "Ráichéal", "Sabah", "Sabeena", "Sabi", "Sabira", "Saj", "Sally-Ann", "Sameera", "Samia", "Sangeet", "Sanjana", "Sasjkia", "Saskia", "Satvir", "Scarlett", "Seema", "Selaine", "Seonaid", "Serena", "Shabana", "Shaghofta", "Shakeela", "Shama", "Shan", "Shanaz", "Shanell", "Shannon", "Sharifah", "Sharmen", "Shazna", "Sheridan", "Sherrie", "Sheryll", "Shivani", "Shâo-Lan", "Sian", "Signe", "Simita", "Simran", "Siobhain", "Siobhán", "Siân", "Smita", "Sonya", "Sorcha-Lucy", "Stacy", "Stefanie", "Steph", "Stephani", "Suella", "Sureena", "Sushila", "Syeda", "Sylvia", "Taghrid", "Taiwo", "Talia", "Tania", "Tanushka", "Tanya", "Tara", "Tara-Jane", "Te", "Thalia", "Thangam", "Thea", "Theresa", "Thirza", "Tilly", "Tina", "Tonia", "Tor", "Tulip", "Uma", "Una", "Veronica", "Vikki", "Virginia", "Vivien", "Vivienne", "Vix", "Wera", "Yolande", "Yuan", "Yvette", "Zara", "Zarah", "Zarmeena", "Zebunisa", "Zia", "Zoë", "Zöe", "Áine", "Órfhlaith", "Catrin", "Jess", "Iris", "Mags")

#below are possibly non-binary, but it is difficult to verify this with confidence

non_binary <- c("RJ", "Athol")

# below are names that cannot be verified

other <- c("A.N.ON", "Adie", "Akshay", "Amer", "Anni", "Baz", "Bex", "Bharatkumar", "Binka", "Dee", "Gwen", "Harl", "Hoz", "Kombat", "Mariatu", "Morven", "Mykel", "Nagbea", "Rayyan", "Sandy", "Shaukat", "Si", "Tamas", "Taukir", "Vaz", "Zephyr")

#matching the names 

gender_check <- name %>%
  mutate(gender = case_when(
    first_name %in% male_names1 ~ "man",
    first_name %in% male_names2 ~ "man",
    first_name %in% male_names3 ~ "man",
    first_name %in% male_names4 ~ "man",
    first_name %in% male_names5 ~ "man",
    first_name %in% male_names6 ~ "man",
    first_name %in% female_names1 ~ "woman",
    first_name %in% female_names2 ~ "woman",
    first_name %in% female_names3 ~ "woman",
    first_name %in% female_names4 ~ "woman",
    first_name %in% non_binary ~ "unknown",
    first_name %in% other ~ "unknown",
    TRUE ~ ""
  ))

gender_check <- gender_check %>%
  select(person_name, first_name, last_name, gender, everything())

#unisex names ordered separately

unisex <- gender_check %>%
  filter(gender == "")

#FEMALE NEUTRAL NAMES

female_neutral1 <- c("Jo Hart", "Alex Baker", "Lesley Ann Taylor", "Jo White", "Tracy Kelly", "Ash Jones", "Jo Bird", "Gita Gordon", "Ali Griffiths", "Alex Fawbert", "Naz Shah", "Gill Leek", "Valerie Gray", "Kerry McCarthy", "Jai Breitnauer", "Sam Ammar", "Jo Stevens", "Meg Shepherd-Foster", "Tracey Hague", "Gill Westcott", "Kay Lauer", "Gill German", "Kay Wesley", "Lesley Backhouse", "Sam Casey-Rerhaye", "Tracey Little", "Alex Mayer", "Tracy Gilbert", "Jo Mowat", "Jo Phillips", "Mel Dawkins", "Pat Cullen", "Lesley Rollings", "Morag McRae", "Meg Hillier", "Sam Lee", "Sammy Barry-Mears", "Jo Robb", "Naz Panju", "Alex Bulat", "Lesley Woodburn", "Nic Cook", "Alix Dearing", "Kerry Elizabeth Harrison", "Sasha Watson", "Jo Platt", "Jean Branch", "Jo Smith", "Chris Northwood", "Kerry Postlewhite", "Rowan McLaughlin", "Alix Mathieson", "Ash Haynes", "Frances Bonney", "Frances Lasok", "Kerry Ann Vosper", "Jean Annlyne Misseys Blackbeard", "Alex Brewer", "Nikki Da Costa", "Alex Davies-Jones", "Gill Furniss", "Sam Banks", "Frances Kneller", "Lesley Veronica", "Bev England", "Alex Sufit", "Leigh Ingham", "Bev White", "Jo Barton", "Sam Goggin", "Jo McKenna")

#MALE NEUTRAL NAMES

male_neutral1 <- c("Charlie Abel", "Graeme Craib", "Kenny MacAskill", "Alex Stevenson", "Chris Philipsborn", "Lee Anderson", "Lee Alan Huntbach", "Chris Nevile", "Robin Millar", "Alex Harrison", "Ray Saint", "Jamie McNamee", "Francis Chubb", "Joe Taylor", "Ryan North", "Ryan Warren", "Gerry Carroll", "Eóin Millar", "Ray Georgeson", "Reggie Adams", "Chris Collin", "Tahir Ali", "Qasim Mohammed Esak", "Lee Dargue", "Rhys Burriss", "Sam Rushworth", "Jamie McGowan", "Chris Webb", "Chris Green", "Joe Salmon", "Alex Burghart", "Robin Tilbrook", "Chris Elmore", "Ashley Fox", "Leigh Redman", "Charlie Dewhirst", "Chris Ward", "Ashley Ridley", "Ray Brady", "Christian Wakeford", "Chris Evans", "Lindsay Whittle", "Jamie Stone", "Lee Canning", "Sam Coates", "Alex Wilson", "Alex Barros-Curtis", "Elliot Colburn", "Ashley Dickenson", "Mel Stride", "Tristan Osborne", "Alex Chalk", "Max Wilkinson", "Chris Chilton", "Chris Brody", "Nic Puntis", "Lindsay Hoyle", "Robin Adamson", "Sasha Jolliffe Yasawi", "Lee Lavery", "Jamie Orange", "Drew Thomas Gilchrist", "Lee Barron", "Lee Forster", "Chris Lofts", "Lee Daniel Gibbs", "Chris Philp", "Sam Holland", "Lee Stranders", "Jamie Mulhall", "Joe Naitta", "Sam Ward", "Chris Williamson", "Lee Pitcher", "Chris Coghlan", "Ashley James Payne", "Chris Tough", "Sam Bastow", "Chris Law", "Naz Anis-Miah", "Ryan Blackadder", "Graeme Downie", "Danny Smith", "Les Beaumont", "Sam Habeeb", "Joe Bhangu", "Tahir Maher", "Danny Donnelly", "Sammy Wilson", "Tahir Mirza", "Chris Scott", "Jai Singh", "Danny Kruger", "Chris Murray", "Alex Martin", "Cameron Rose", "Lee Evans", "Charlie Davis", "Sam Gabriel", "Christian Hacking", "Ryan Coogan", "Chris Dey", "Lee Bunker", "Toni Giugliano", "Alex Just", "Alex Deane", "Chris McFarling", "Robin Moss", "Pat O'Connor", "Chris Sermanni", "Danny Raja", "Hal Hooberman", "Chris Farmer")

male_neutral2 <- c("Alex McIntyre", "Kris Callander", "Chris Evans", "Chris Annous", "Sam Peters", "Alex Ballinger", "Ryan Ashley Priest", "Robin Lambert", "Chris Noyce", "Alex Diner", "Alex Cunliffe", "Noel Willcox", "Joe Emmett", "Joe Morris", "Chris Furlong", "Chris Lucas", "Charlie Clinton", "Sunny Brar", "Chris Burnicle", "Sam David Woods-Brass", "Alex Wilson", "Jas Athwal", "Chris McEleny", "Drew Hendry", "Joe Robertson", "Cameron Palin", "Max Jefferson Nelson", "Terry Stacy", "Jamie Rickelton", "Chris Adams", "Joe Powell", "Ali Abdulla", "Rowan Adam Halstead", "Chris Foren", "Chris Howley", "Alex Sobel", "Reggie Harry Wray", "Sam Firth", "Chris Whiteside", "Chris Bell", "Lee Anthony Farmer", "Max Chauhan", "Danny Sweeney", "Danny Bowman", "Sam Gorst", "Joe Owens", "Cameron Glasgow", "Charlie Evans", "Chris Passmore", "Sam Easterby-Smith", "Sunny Virk", "Francis Whitefoot", "Mo Waqas", "Chris Curtis", "Max Windsor-Peplow", "Jamie Dobson", "Lee Raymond James Dillon", "Ali Avaei", "Ashton Hektor Muncaster", "Mad Hatter", "Sam Nicholson", "Chris Hilditch", "Ashton Victor Howick", "Tristan Morrow", "Ash Leaning", "Alex Easton", "Chris Bradburn", "Sunny Moon-Schott", "Lee Rowley", "Alex Zychowski", "Chris Hinchliff", "Ash Cartman", "Sam Carling", "Elliot Alister Tong", "Noel Matthews", "Jamie Osborn", "Sam Harvey", "Alex Norris", "Sam Al-Hamdani", "Alex Armitage", "Graeme Casey", "Chris Goodall", "Alex Kerr", "Alex Moore", "Jamie Luke Needle", "Joe Biddulph", "Joe Cronin", "Charlie Murphy", "Lee Jamie Roberts", "Chris Taylor", "Chris Jones", "Chris Bryant", "Qasim Ajmi", "Lee Martin Taylor", "Chris French", "Danny Clarke", "Chris Copeman")

male_neutral3 <- c("Chris Rowe", "Chris Clowes", "Joe Wood", "Chris Harwood", "Lee Francis Derrick", "Christian Vassie", "Aaron Jacob", "Sam Christmas", "Sam Chapman", "Lee Waters", "Waqas Ali Khan", "Alex Wagner", "Mad Mike Young", "Ryan Kett", "Robin Jackson", "Chandra Sekhar Muvvala", "Max McLoughlin", "Mel Lucas", "Robin Swann", "Chris Carter-Chapman", "Chris Twells", "Chris Hazzard", "Lee John Clark", "Rhys Rhobat Baker", "Chris Brown", "Chris Harrison", "Terry Jermy", "Alex Clarkson", "Joe Greenhalgh", "Pat Moloney", "Chris Kane", "Ashley Walker", "Chris McDonald", "Joe Dancey", "Chandra Kanneganti", "Sam Harper-Wallis", "Noel Sands", "Joe Hudson-Small", "Chris Clarkson", "Chris Eynon", "Ryan Powell", "Les Willis", "Ryan Jude", "Ryan Trower", "Cameron Thomas", "Pat McCarthy")

# UNKNOWN

unknown_neutral <- c("Terry Robinson", "Jamie Pullin", "Chris Moore", "Jamie Alan Hinton-Wardle", "Charlie Caiger", "Kris Stone", "Chris Baddon", "Robin Burnham", "Gerry Cullen", "Lee Moffitt", "Ryan Ahmad", "Ray Bolster", "Chris Whaley", "Chris Ward Nicholson", "Ray James", "Chris Carter", "Ali Khan", "Chris Johnston", "Mel Todd", "Rowan Fitton", "Chris Bovill", "Chris Magness")

# Assign gender based on the name vectors
gender_check$gender <- ifelse(
  gender_check$person_name %in% male_neutral1, "man",
  ifelse(gender_check$person_name %in% female_neutral1, "woman",
         ifelse(gender_check$person_name %in% unknown_neutral, 
                                  "unknown", gender_check$gender)))


#adding in the YouGov poll of most likely to get the seat


```

